\section{The problem}
\label{sec:problem}

Something shifts, speakers of another language arrive in numbers, a new medium spreads, a border is redrawn, a government issues a decree, and a form that was rare before it turns common after. Whether or not the shock caused the change is a fair question, and corpus linguistics is well equipped to ask it: the before and the after are measurable, and the field has long studied how change actually propagates \citep{weinreich1968,tagliamontedarcy2007}. The causal reading still enters easily, often through a verb, usage that \enquote{responds to} a reform rather than an argued claim. Corpus work on prescription has long felt this pull: confronting normative dictates with usage, it tends to find the causal link assumed rather than demonstrated, from relativizer choice in edited English prose \citep{HinrichsSzmrecsanyiBohmann2015} to five centuries of French injunction set against the record of speech \citep{PoplackDion2009}. The cheapest robustness check is sometimes grammatical: delete the causal verb and see what claim is left. A rise after a shock is a correlation in time, and what turns it into a causal one is a discipline the corpus makes harder, not because linguists neglect it but because a corpus adds confounds of its own that a plain before-and-after comparison cannot separate from real change.

In econometrics, the standard repair is difference-in-differences (DiD): set the treated series against a comparison series the shock left untouched, and read the effect off the change in their gap. The design isn't new to language research, where causal inference from text is itself an active field \citep{keith2020text,grimmer2022text,egami2022texts}, but it's nobody's default for tracking change in a corpus. That work has mostly handled text as a treatment, outcome, or confounder in a cross-section; less settled is what DiD's own assumptions, parallel trends and no interference, become for a diachronic corpus frequency.

The frequency is a constructed aggregate, built from many texts by choices about what to count and how to weight, the same shape as a price index, so the discipline it calls for is not new; what's new is bringing it to corpus linguistics, where it's rarely applied, and to a filter harder than most, since who gets written about and how a form is coded are read off the same text and seldom measured. Carrying it across is the gap this paper works in.

A corpus is not simply a sample of a population's usage; it's a sample of texts, assembled by editors and archivists with their own shifting priorities \citep{biber1993}, and its balance and representativeness are matters of degree that builders approximate rather than attain \citep{McEneryHardie2012}. Tracking change compounds this: the matched diachronic corpora built for it, like the Brown family, must keep a sampling frame comparable while the genres inside it drift \citep{LeechHundtMairSmith2009}. So a corpus frequency carries two things at once, what the population did and who happened to be in the corpus, and a shock can move either. The DiD machinery is sound; what it needs is to be told which of the two it's identifying.

The contribution is therefore not a tutorial on running DiD but a discipline for claiming with it. A before-and-after corpus contrast supports a causal reading only after several things are pulled apart: the target population, the corpus composition, the measurement procedure, the unit at which the treatment varies, and the unit at which uncertainty is counted. Conflate any two and the apparent effect can be an artifact of the corpus, not a fact about the language.

Six questions organize the discipline, taken up across the sections that follow: what's the shock, and which act counts as the treatment (Section~\ref{sec:treatments}); at what unit does the treatment actually vary (Section~\ref{sec:estimand}); what population's rate is the target, and what outcome measures it (Sections~\ref{sec:estimand} and~\ref{sec:outcomes}); what comparison makes the counterfactual plausible, and what corpus-specific threat could counterfeit the effect (Section~\ref{sec:threats}); and what diagnostic would expose it (Section~\ref{sec:diagnostics}). Section~\ref{sec:modern} takes up the estimation and inference choices the answers force.

Corpus coverage lags the method's reach, though less than one might fear. It skews toward well-documented written traditions, but not narrowly: large historical corpora exist for many European languages, for Chinese, and for others, so the before-and-after series the design needs run well past the modern standardized case. Where none exists the design falls silent, and which languages get a corpus at all is the composition problem at its largest scale.

A single case runs through the paper to keep the questions concrete: the feminization of profession nouns under francophone language policy, a change tracked in corpus studies of French \citep{Fujimura2005,BurnettBonami2019}. It earns its place in Section~\ref{sec:worked} as a teaching vehicle, not as a finding; the aim is the method, and the case is built so that its verdict is sometimes that the effect cannot be identified. A corpus DiD is worth most when it disciplines a causal claim downward, and that's the argument the rest of the paper makes.
