\section{The DiD estimand in corpus terms}
\label{sec:estimand}

DiD asks one question: across a shock, did the treated series and a comparison series no shock touched change differently? Take the treated variety's before-to-after change in the outcome, subtract the control's, and under a single assumption (that absent the shock the two would have moved in parallel) the remainder estimates the average treatment effect on the treated (ATT) \citep{RothSantAnnaBilinskiPoe2023}. Say feminine marking rose ten points in the treated variety across the window and four in the comparison: DiD reads the policy effect as the six-point excess, on the assumption that without the policy the treated variety would have risen by the same four. Everything corpus-specific lives in what fills \enquote{outcome} and \enquote{series}.

That last clause is where the textbook panel idealization, the clean grid of the same units measured period after period, breaks down for a corpus as for any constructed aggregate. What a corpus yields is an observed frequency $f$; the quantity of interest, the estimand, is the latent rate $\pi$ at which a specified population uses the form in a specified context, for the running case the share of profession- and title-noun tokens with a female referent that take feminine marking, in a named text population. Both the population and the context belong to the estimand: fix them too late and one composition fact reads as nuisance bias in one rung and as part of the target in the next. The two differ by everything the corpus adds: write $f = \pi + b$, with $b$ collecting composition (whose texts, which genres, what topics) and measurement (how the form was coded or transcribed).

Feminine marking can climb in the corpus with no change in anyone's usage at all: a paper that runs more profiles of women in the professions, or a transcription service that starts spelling out feminine titles it used to abbreviate, moves $f$ while $\pi$ holds still. Read $b$ as the gap itself, $b = f - \pi$: not a random error that more data averages away, but whatever composition and measurement added, fixed once the texts and the coding are. So $f = \pi + b$ holds by definition, and collecting more tokens shrinks the sampling noise around an estimate without touching $b$. That sampling noise is a separate quantity, carried by the cell-level uncertainty of Section~\ref{sec:modern}. The outcome on the page is not the quantity the claim is about.

The residual is not empty. Stratify the texts by the margins composition moves on (outlet, profession, topic), and within each stratum $s$ let $\pi_s$ be the true rate, $r_s$ the rate the corpus records, $q_s$ the corpus's share of $s$, and $p_s$ the target population's. Then $f = \sum_s q_s r_s$ and $\pi = \sum_s p_s \pi_s$, so
\[
b = \underbrace{\sum_s (q_s - p_s)\,\pi_s}_{\text{composition}} + \underbrace{\sum_s q_s\,(r_s - \pi_s)}_{\text{measurement}}.
\]
Reweighting sets $q_s$ to $p_s$ on the strata it can see, zeroing the first term; the second survives, and the part of it the strata are too coarse to resolve is the hidden drift of Section~\ref{sec:worked}. The simulations are built on this split.

Differencing is linear, so the DiD of $f$ separates into the DiD of $\pi$ and that of $b$, $\mathrm{DiD}(f) = \mathrm{DiD}(\pi) + \mathrm{DiD}(b)$. The first term, under parallel trends in the rate, is the target effect. For a usage-rate estimand, the second is contamination, and it vanishes only under a second parallel-trends assumption, $\mathrm{DiD}(b) = 0$: that composition and measurement would have drifted alike in both series. A corpus design needs parallel trends twice over, in $\pi$ and in $b$, where the textbook case quietly assumes the second away. Picture one variety's archive getting re-digitized and re-tagged halfway through the window while the other's is left alone: the two corpora's coding has now drifted apart even if actual usage tracked perfectly, and that drift would enter the estimate as if it were the policy's effect.

Both are claims about an unobserved counterfactual: pre-treatment comparisons can discredit them but cannot verify them. And since only $f$ is observed, a flat pre-period in $f$ is consistent with non-parallel trends in $\pi$ and in $b$ that cancel, so the pre-period tests the two assumptions jointly, never apart.

Neither half is optional, and the temptations run both ways. Once the composition term is on the table it's easy to treat it as the whole problem and forget the ordinary assumption about the rate; it's equally easy to wave at adjusting for composition and assume the rate takes care of itself. Both shortcuts fail. The observed contrast recovers the effect on $\pi$ only when the rate would have moved in parallel \emph{and} the bias would have moved in parallel.

One refinement decides how the corpus-filter term is handled, turning on which effect is wanted. The filter reaches the observed contrast two ways: a non-policy drift, which the second assumption requires to be parallel, and a policy-induced shift, the policy changing what gets written rather than how a given referent is marked. For an effect on the usage rate $\pi$, that second shift is nuisance as well, even though the policy caused it: a topic or profession mix that moved gets reweighted to a fixed mix, isolating the marking rate. The other estimand, sometimes the one actually wanted, is the policy's total effect on the observed output. It keeps the writing that the policy moved, so that same shift is part of the answer, and reweighting it away would subtract a piece.

The two estimands differ exactly by the policy-to-composition path, and a design that doesn't say which it means can defend neither. The one margin neither estimand touches is the outcome's own, the realization strategy of Section~\ref{sec:outcomes}: condition there and the effect itself is removed.

Naming the estimand forces apart five units a corpus invites running together. The \term{linguistic unit} is the variable whose rate is at issue (feminine marking of a profession noun for a female referent). The \term{measurement unit} is what gets counted to estimate that rate (eligible tokens, or documents). The \term{aggregation unit} is the cell those counts are summed into before estimation (an outlet-month, a title-variety-year), which sets the weights and denominators and can quietly change the estimand. The \term{identifying unit} is where the treatment varies (the variety or polity). The \term{dependence unit} is the level at which residuals correlate and inference is clustered, here the variety again; with only a few varieties the usual large-sample standard errors don't hold, which is the inference problem Section~\ref{sec:modern} takes up.

The merge is dangerous because the five sit at wildly different scales. A corpus offers millions of measurement units and, often, two or three identifying units, and it's the latter that govern how much can be concluded. Counting tokens as though they were the sample size manufactures a precision the design doesn't have \citep{BertrandDufloMullainathan2004}: it's like claiming a hairline margin of error on a two-person poll because each person answered a thousand questions. The estimand is the effect on $\pi$; parallel trends in $\pi$ and in $b$ is its price; the diagnostics of Section~\ref{sec:diagnostics} can falsify that assumption or bound how far it may fail, never confirm it, and the worked example of Section~\ref{sec:worked} is where the discipline is applied end to end, each rung declaring its estimand in a ledger (Table~\ref{tab:ledger}).
