\section{What counts as treatment in language data}
\label{sec:treatments}

In a policy evaluation the treatment is usually a known act on a known date: a law takes effect, a price changes, a program opens. Language data rarely offers that. A change in usage is pushed by a tangle of events, a circular, a dictionary, an administrative endorsement, a first enforcement, a media house revising its style sheet, and which of them counts as the treatment is not given by the data; it's a decision that fixes the estimand. Name the wrong act and the cleanest design answers the wrong question.

And the shock need not be a policy at all. It can be a new medium with constraints of its own (a platform's character limit, a switch to keyboards with autocorrect), a contact or demographic event (a border redrawn, a refugee influx), a censorship regime imposed or lifted, or a non-linguistic event that drags vocabulary behind it (a pandemic, a new technology). These often make for cleaner identification than a decree does, because their timing isn't chosen by the speech community: a platform changes its limit, or a population is displaced, on a schedule unrelated to the linguistic trend, where a policy's timing is bound up with the very trend at issue (Section~\ref{sec:threats}). The policy case this paper runs on is, in that sense, the demanding one for identification.

Take the feminization case. The candidate treatments for one polity include the legislative or administrative endorsement, the publication of a feminized-titles dictionary \citep{Moreau1991}, the issuing of a circular, and the date editors actually began applying the forms. These differ in time, sometimes by years, and in force. A design that dates the treatment to the circular and one that dates it to editorial uptake estimate different effects, and only one of them is the effect being claimed.

The widest gap is between announcement and implementation. A rule can be promulgated and then ignored: France's 1986 circular \citep{france1986circ} urged feminine forms and went largely unenforced, by the government's own later admission, until a second circular in 1998 \citep{france1998circ,BurnettBonami2019}. The two share their text and differ in force, which is what makes 1986 usable as a near-null treatment, a policy on paper whose null result is a negative control, a case that should show nothing, not a failure.

Announcements can also lag the changes they name. An endorsement often ratifies a shift already under way rather than starting it, a question of actuation, of why a change takes hold when and where it does \citep{BurnettBonami2019}, so the official date can be a waypoint on a rising curve, not its onset, and a jump measured there reflects anticipation, not effect. The pre-period diagnostic of Section~\ref{sec:diagnostics} tests for exactly that; here it's a reason to define the treatment as a change in the rate of adoption, not the first appearance of a form.

A deeper precondition sits on the change, not just the policy. Feminization is a gradual, socially structured diffusion, an S-curve that spreads token by token and register by register \citep{weinreich1968,tagliamontedarcy2007}. Two varieties climbing parallel S-curves still satisfy parallel trends, equal nonzero slopes are fine; the threat is their being out of phase. Say one variety sits in the steep middle of its curve while the other is near the top, leveling off. With no policy at all their slopes differ, because an S-curve's slope tracks $\pi(1-\pi)$: a variety at the half-way mark climbs almost three times as fast as one at nine-tenths ($0.25$ against $0.09$). Where the treated and control varieties sit at different points on their own curves, their untreated slopes differ and the counterfactual is nonlinear, so a linear pre-trend extrapolation mis-predicts it and books the curvature as effect, policy or no policy.

A shock identifies an effect only when it shows as a level shift or an acceleration on top of curves that were running parallel; where the varieties are out of phase, \enquote{not identified} becomes the honest modal verdict by the geometry of the change, not an accident of thin data. The cleanest cases for a usage effect are genuinely discrete shocks like a hard enforcement date or an orthographic reform; a transcription convention that flips is discrete too, but it's a clean measurement shock, a rival event to diagnose, not a usage effect.

Treatments in this domain are rarely binary either. Endorsements vary in scope (which titles, which registers) and in teeth (recommended versus required), so two polities labelled treated may have had very different doses. Where the dose is measurable, an intensity rather than an on-off indicator is the more honest treatment; where it isn't, the binary coding is itself an assumption to flag.

Finally, the treatment's boundary leaks. A policy in one polity reaches readers in another through shared media and the channels a common language opens (Section~\ref{sec:threats}), so a variety marked untreated can receive a dose of the treatment it's meant to control for. The treatment is not only fuzzy in date and force but porous across units, which is why naming it has to include naming its edges.

So the first question of a corpus DiD is not yet about trends or controls; it's which act, at which date, with what reach, is being called the cause. The estimand, the comparison, and every diagnostic that follows inherit that choice, and leaving it implicit is how a study ends up defending a number whose meaning it never fixed.
