\section{Outcome construction}
\label{sec:outcomes}

A corpus outcome is transparent when anyone could reproduce it from the same texts by a stated rule: a count over forms, a closed-set alternation, a tagged feature. It's model-dependent when a classifier, parser, or embedding stands between the text and the number. Both are usable, but they fail differently. A transparent count errs in what its rule misses; a model-dependent measurement adds the model's own drift, which can move with the very period or variety under study and so mimic the effect. That makes the check differential, not global: a classifier or parser has to be evaluated by period and by variety, since an error rate that holds on average but shifts across the window is the measurement twin of a non-parallel trend.

And because the outcome is a proportion, the bias is level-dependent even at fixed error rates: it tracks $\pi$ itself, so a moving rate drags the filter through the measurement model, coupling the two parallel-trends assumptions in mechanism, not only in the data. A classifier that mislabels one token in twenty, both ways, doesn't add a flat one-in-twenty bias to the share: a true rate of ten percent reads as fourteen, a true ninety as eighty-six, and a true fifty as fifty exactly, the distortion flipping sign as the feature grows common, all from a classifier that never changed.

A second choice hides inside the first. Most variables of interest are realized in more than one way, and a single count silently treats those ways as exchangeable signs of one change. Feminine marking of a profession noun for a woman, this paper's running example, surfaces as a distinct noun form (\mention{autrice}), as a feminine article or agreement on an otherwise invariant noun (\mention{la ministre}), or as both. Pooling them assumes the realizations rise together; breaking the outcome out by strategy tests that assumption instead of burying it, and the test matters because varieties favour different strategies \citep{Dawes2003}.

Two refinements follow. Agreement that's already obligatory sits at a ceiling: agreement on an already-feminine title (\mention{une infirmi{\`e}re comp{\'e}tente}) varies with nothing the shock touches, so it's noise, not outcome, and the informative agreement is the contested kind, the feminine article on a default-masculine title, which for epicene nouns (one form for either gender, like \mention{ministre}) is the only available signal. And a marker means nothing until its conditioning is fixed: a feminine article counts as feminization only when the referent is independently known to be a woman, so the outcome definition carries that conditioning rather than assuming it.

The outcome is a proportion, and a proportion needs a denominator named as carefully as the numerator. Feminine-marked tokens out of what: every profession-noun token, only those with a female referent, only title-address contexts? Weighted how: by token, so a few prolific titles dominate, or by title-variety cell, so each counts once? Over which referent distribution, when the share of female referents is itself moving? The numerator is the visible choice; the denominator is where a composition shift quietly re-enters, which is why it belongs in the estimand of Section~\ref{sec:estimand}, not in the counting code.
