\section{Identification threats specific to corpora}
\label{sec:threats}

The threats a corpus difference-in-differences faces are easiest to read off a small set of causal paths (Figure~\ref{fig:threat-dag}). They fall into five families. Four are inflected, or created, by the corpus, composition, measurement, interference, and aggregation; the fifth, selection into treatment timing, is the ordinary confounding worry any difference-in-differences faces. Each is a way the observed contrast can move while the population rate stays put, and each has a diagnostic in Section~\ref{sec:diagnostics} that can expose it.

\begin{figure}[t]
\centering
\begin{tikzpicture}[>={Stealth[length=2mm]}, every node/.style={font=\small, align=center}]
  % top row: ideology and the children that reach f without passing through pi
  \node[draw, dashed, ellipse, inner sep=1pt] (I)  at (1.6,1.8)   {Ideology};
  \node[draw, rounded corners]                (Tn) at (4.6,1.8)   {Neighbour policy};
  \node[draw, rounded corners]                (M)  at (7.6,1.8)   {Coding /\\transcription};
  % middle row: the causal spine
  \node[draw, rounded corners]                (T)  at (0.8,0)     {Policy timing};
  \node[draw, dashed, ellipse, inner sep=1pt] (pi) at (4.0,0)     {Pop.\ rate $\pi$};
  \node[draw, thick, rounded corners]         (f)  at (8.4,0)     {Observed $f$};
  % bottom row: the composition layer
  \node[draw, rounded corners]                (Sc) at (1.2,-1.8)  {Non-policy\\composition};
  \node[draw, rounded corners]                (C)  at (4.0,-1.8)  {Composition $C$};
  \draw[->] (I) -- (T);
  \draw[->] (I) -- (pi);
  \draw[->] (I) -- (Tn);
  \draw[->] (I) to[bend left=30] (M);
  \draw[->] (T) -- node[above, font=\scriptsize] {effect} (pi);
  \draw[->] (T) -- node[below left, font=\scriptsize, pos=0.6] {mediator} (C);
  \draw[->] (Sc) -- (C);
  \draw[->] (pi) -- (f);
  \draw[->] (C) -- (f);
  \draw[->] (M) -- (f);
  \draw[->, dashed] (Tn) -- node[above right, font=\scriptsize, pos=0.2] {spillover} (f);
\end{tikzpicture}
\caption{The causal paths a corpus difference-in-differences must contend with: a threat map, not a graph that identifies the effect. Dashed nodes are unobserved (gender ideology and the population rate $\pi$). The arrows are the routes the section takes up in turn, ideology as a common cause of adoption timing and usage, the policy's effect on the rate (the target) and its mediated effect on composition $C$, the measurement link from $\pi$ to observed $f$, and a neighbour's spillover. What the figure cannot draw is the assumption that identifies the effect: parallel trends is a restriction on a time-trend, not a path a static graph can show.}
\label{fig:threat-dag}
\end{figure}

Composition is the first, and it splits in two. A policy can move the very mix it's measured against, the professions in the news, the topics in play, so part of composition lies on the path from treatment to $f$ (the edge $T \to C$) and is a mediator: reweighting on it subtracts part of the effect. The rest, sampling frame, outlet entry, author turnover, register mix, drifts for reasons the policy didn't cause, and that part is a bias source to adjust for. The two look alike in the data and demand opposite handling, which is why the margins to reweight have to be named before the estimate is seen.

Measurement is the second. What the corpus records is not the latent rate but a coding of it, and the coding can drift, a classifier whose error rate moves with the period, a transcription house-style that adopts new forms on a date of its own. When it does, $f$ jumps while $\pi$ holds still. That's confounding read off the text itself \citep{keith2020text}, the instrument measuring the outcome turned into a moving part, and it bites hardest where the same ideology that drives usage can also revise the coding convention.

Interference is the third, and it breaks the assumption that one variety's treatment leaves the others untouched. Suisse romande reads French media, both draw on the same wire service, and a policy in one polity reaches readers in another. A not-yet-treated control is then not untreated but lightly dosed, so the estimand has to be named, the policy's direct effect net of diffusion, or its total effect including it, and the shared material held aside for a separate estimate.

Aggregation is the fourth, and the easiest to overlook. The level at which counts are summed into observations, token, document, outlet-month, title-variety-year, fixes the weights and the denominators, and a change in that level can move the estimand with no change in usage: a few prolific outlets or a burst of one topic can carry a token-level rate that a cell-level rate would not. The aggregation unit of Section~\ref{sec:estimand} is a modelling choice with identifying consequences, not bookkeeping.

Selection into treatment timing is the fifth and the deepest. A polity tends to legislate about when its gender politics have already moved, and that same movement shifts usage, so adoption timing is correlated with the outcome's own trajectory. The ideology fork in the figure is exactly that, a common cause of when the policy lands and of how usage was already going. No estimator removes it, and in a two-group design it's observationally indistinguishable from a real effect.

That last threat is also where the figure reaches its limit. The fork it draws is a static object, but identification turns on a trend, whether usage would have moved in parallel absent the policy, in the rate and in the filter both. A graph can inventory the threats; it can't draw the assumption that answers them, which is why the design differences the confounding out rather than conditioning on it, and why the diagnostics of Section~\ref{sec:diagnostics}, not the diagram, carry the weight.
