\section{Diagnostics}
\label{sec:diagnostics}

A diagnostic here does two things and not a third. It can disqualify a causal reading, or bound how far the identifying assumption may bend before the result flips. It cannot confirm parallel trends, since the counterfactual it concerns is never observed. So diagnostics are not a checklist run after the estimate to bless it; they're pre-specified tests, each answering a threat from Section~\ref{sec:threats}, whose outcomes route the estimate to one of a few permitted conclusions under a rule fixed before the estimate is seen.

That discipline starts before any test, with the roles. Each textual feature has to be assigned a job: does it \emph{measure} the outcome, \emph{identify} the referent that conditions the outcome, \emph{index} the composition to adjust for, or lie \emph{downstream} of usage as part of the outcome itself? No feature can hold two of these jobs at once. Conditioning on a downstream feature, the realization strategy, subtracts the effect; conditioning on a measurement cue that drifts imports the drift; and how far composition is reweighted follows from the declared estimand (Section~\ref{sec:estimand}). Writing the roles down in advance is what keeps the later adjustments honest.

The first family tests the pre-period. Event-study estimates against not-yet-treated comparisons \citep{SunAbraham2021,BorusyakJaravelSpiess2024} show whether the series moved together before the shock, but a flat pre-period only fails to discredit the assumption; it cannot establish it \citep{RothSantAnnaBilinskiPoe2023}. The sharper question is onset versus acceleration: does the adoption rate change at the treatment date, or was the divergence already under way, so the date is a waypoint on a curve rather than its cause? A slope already bending before the shock routes away from a causal reading.

The second family breaks the design on purpose. A placebo date where no policy occurred, a near-null policy used as a negative control, or the treatment label permuted across units should all return nothing, and an effect where none belongs discredits the pipeline. With a handful of identifying units these permutations are calibration, not inference: they show the size of contrast the procedure manufactures by chance, against which the real one is read, not a $p$-value from an assignment that was never random \citep{BertrandDufloMullainathan2004}. What inference remains possible with so few clusters is the business of Section~\ref{sec:modern}.

The third family withholds parts of the corpus. Leave-one-source-out re-estimation, dropping an outlet, a wire feed, or a region, asks whether the effect lives in the whole series or in one corner of it, and leave-one-author-out belongs beside it, since a few prolific bylines or a single topic burst can carry a token-level rate that a cell-level rate would not; syndicated reprints of one article are deduplicated first. An estimate carried by a single outlet, a single author, or shared wire copy is idiosyncrasy or interference, not an effect.

The fourth family probes the corpus filter. Reweighting to a fixed composition removes a measured artifact; which margins to fix follows from the estimand (Section~\ref{sec:estimand}), the policy-moved mix reweighted away for a usage effect and kept for a total-output effect, the choice pre-committed either way. And because the outcome is a bundle of realization strategies (Section~\ref{sec:outcomes}), re-running it strategy by strategy is itself a diagnostic: the strategies sit on a cline, low-cost agreement (\mention{la ministre}) moving early and high-cost derivation (\mention{autrice}) late, so a per-strategy event study says not just whether a policy bit but which strategy it moved, and a divergence between them means the assumption that they're exchangeable signs of one change has failed.

The last family quantifies the assumption's fragility. Rather than assert parallel trends, sensitivity analysis asks how large a violation would have to be to overturn the result, and compares that breakdown value to the largest confound the setting plausibly supplies \citep{RambachanRoth2023}. To keep \enquote{plausibly} from meaning \enquote{whatever the result needs}, the benchmark is fixed in advance and measured, not felt: the largest trend swing seen in an untreated stretch or an untreated variety sets the bar a real violation would have to clear, and the margins eligible for reweighting are listed from domain knowledge before the estimate is in hand. A result that survives only a violation smaller than that benchmark has not earned its claim; one that survives a larger violation earns a bounded claim, reported as a sensitivity interval, not a point. The benchmark has a blind spot worth naming: it can only see a violation that leaves a footprint in the windows it's built from. A drift confined to the treated post-period, or one large enough to inflate the benchmark along with the estimate, clears the bar untouched, which is exactly the failure the simulations of Section~\ref{sec:worked} make quantitative. The bound is eliminative: it clears the confounds you can name and benchmark, never the ones you cannot conceive, so a bounded claim is always made against the imaginable confounds, not all the possible ones.

These checks feed one pre-specified rule, of the same form in every application:
\begin{itemize}
\item a divergence aligned to the treatment date that survives the composition, interference, and onset checks, with a breakdown value beyond any plausible confound, licenses a \emph{bounded} causal estimate, scoped to the register actually measured;
\item a divergence that leans on an ambiguous check, or whose breakdown value sits near a plausible confound, licenses only a \emph{descriptive} reading;
\item calendar-time co-movement with no date-aligned divergence is read as a \emph{shared wave}, a secular trend rather than the intervention, an identified result in its own right;
\item a pre-treatment divergence, an effect carried by shared-source or single-outlet material, selection into treatment timing the design cannot rule out, or a breakdown value below a plausible confound, forces \emph{not identified}, reported as such rather than called robust.
\end{itemize}

The defeater itself needs discipline. An unmeasured local wave is always conceivable, but a verdict of \enquote{not identified} resting only on what one can imagine, never on what the setting gives reason to suspect, would collapse the ladder to a single rung; the defeater has to be credible in the case at hand, not just conceivable, or no case ever clears the bar.

The ladder's point is that a downgrade is a result. A corpus DiD succeeds not when it returns an effect but when it returns the right verdict, and for many questions that verdict is that the apparent change is a shared trend, a measurement shift, or a composition artifact, or else that no usage effect can be identified from the series at all. The worked example of Section~\ref{sec:worked} runs the ladder on one case; that the rule was fixed before the data answered matters more than which rung the case lands on.
