\section{Estimation and inference choices}
\label{sec:modern}

An estimator identifies nothing on its own; the design does. The choices in this section, which estimator, on what scale, with what inference, cannot rescue a contrast the design fails to identify, but a careless one can squander or misstate what the design does buy. That's why the section is about estimation and inference, not only about which DiD package to call.

Start with the estimator, because the default is a trap. A pooled two-way fixed-effects regression looks like the natural way to absorb variety and period, but under staggered adoption with heterogeneous effects its implied weights can turn negative, so already-treated units serve as controls for later ones \citep{GoodmanBacon2021,deChaisemartin2020}. The group-time and event-study estimators built for this case repair the weighting \citep{CallawaySantAnna2021,SunAbraham2021,BorusyakJaravelSpiess2024}. They repair only the weighting, though: none of them touches selection into treatment timing, so a heterogeneity-robust estimate of an endogenously timed treatment is a clean number for the wrong quantity.

The estimand also lives on a scale, and the scale is a modelling choice. The additive split $f = \pi + b$ buys clean linearity, but parallel trends is not transform-invariant: an assumption that holds on the probability scale can fail on the log-odds scale and the other way round, and the gap between them is widest when the varieties start from different baseline levels \citep{RothSantAnna2023}. A logit-scale parallel-trends may be the more defensible assumption, but on that scale the tidy additive decomposition no longer carries through, so the choice is substantive, not cosmetic. Binomial or beta-binomial models estimate cell rates efficiently without changing what's identified; whatever the fitting scale, report the effect back on the scale the claim is made.

Inference is where the few-units problem bites hardest. The binding sample size is the number of identifying units, not the millions of tokens, and with four or five polities cluster-robust intervals are badly anti-conservative \citep{BertrandDufloMullainathan2004}. The treatments that exist for this regime are the wild cluster bootstrap \citep{CameronGelbachMiller2008} and randomization inference for designs with few treated clusters \citep{MacKinnonWebb2020}. Neither is a guarantee: the bootstrap can still under-cover with very few or very unbalanced clusters, and randomization inference needs a credible assignment model that staggered, politically chosen adoption rarely supplies. They sharpen the honesty of the interval; they do not manufacture one. With as few as two identifying units, or with politically chosen treatment timing, they may be more illustrative than runnable: where their assumptions can be made explicit, run them; where they can't, report the absence of a clean interval as part of the result. And the unit of analysis should be pre-specified cells, an outlet-period or a title-strategy cell, not raw tokens treated as independent draws, which is the aggregation choice of Section~\ref{sec:estimand} seen from the inference side.

This invites the obvious objection: why difference at all, rather than model the whole series, a multilevel model with latent rates, outlet and title effects, composition terms, and a smooth calendar trend? Not because differencing removes more. Differencing clears the time-invariant confounds, the fixed level of a variety's gender politics, but it leaves the trend confound exactly where a model would: selection into treatment timing rides the trend, and no estimator, differenced or modelled, removes it. The identifying assumption is the same one either way, parallel trends here, a counterfactual-trend term there, and a pre-period checks it equally for both. What DiD buys is not a stronger guarantee but a named and conventional one: the assumption is called parallel trends, it sits in plain view, and a flat pre-period is its agreed probe, where a multilevel model can bury the same assumption in functional form and the random-effects structure. Partial pooling is still a good estimator for the nuisance dimensions, shrinking sparse outlet and title cells within each variety while the across-variety contrast stays fixed (Section~\ref{sec:worked}); it's a complement for estimation, not a substitute for the design's identification.

None of these choices identifies anything. Each can only protect or squander what the parallel-trends design already bought, which is why they belong after the estimand and the diagnostics, and before the worked example puts them to work.
