\section{Worked example}
\label{sec:worked}
\FloatBarrier

I build the design up rather than show it whole. Each rung adds one source of difficulty and turns the threat it introduces into a lesson, so the reader can see what a corpus DiD identifies and where it stops. The phenomenon throughout is the feminization of profession nouns under language policy.

Each rung declares its estimand before any estimate (Table~\ref{tab:ledger}); the difference that matters most is the target, the usage rate $\pi$ in rungs 1 and 2 and a design contrast rather than a clean effect in rung 3.

\begin{table}[t]
\centering
\footnotesize
\caption{The estimand ledger. Each rung fixes its target, design, and effect type before any estimate. Rung 3 is a design variant, not a clean two-group contrast.}
\label{tab:ledger}
\begin{tabular}{@{}l p{0.235\textwidth} p{0.235\textwidth} p{0.235\textwidth}@{}}
\toprule
 & \textbf{Rung 1} & \textbf{Rung 2} & \textbf{Rung 3} \\
\midrule
Target          & usage rate $\pi$, edited written standard & usage rate $\pi$, edited written standard & language-internal vs polity-driven (a design contrast) \\
Treatment       & Swiss 1991 endorsement & staggered endorsements, 1979--1998 & no within-Switzerland contrast; federal policy fixed \\
Comparison      & France (1986 near-null) & not-yet-treated polities & other language regions (not untreated controls) \\
Window          & 1991--1997 & 1979--1998 and after & around 1991 \\
Outcome         & feminine-marking share, female-referent profession/title tokens & the same, strategy-decomposed & language-specific marking, harmonized as far as possible (\emph{Bulletin officiel}) \\
Identifying unit & variety & polity & language region within one polity \\
Dependence unit & variety (two) & polity (four) & region (three) \\
Effect          & direct or total, declared & group-time ATTs & not a clean DiD; exposes the measurement link \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Rung 1: the clean contrast and the spillover problem}

The cleanest start is a pair of varieties that share a language and differ in policy. French-speaking Switzerland endorsed feminizing profession nouns early: the Swiss Federal Chancellery's 1991 report on non-sexist administrative language \citep{chancellerie1991} and \textcite{Moreau1991}'s professions dictionary. France makes the comparison, but not a pristine control. A 1986 circular under Prime Minister Fabius urged the same forms yet went largely unenforced, by the government's own later admission, until a second circular under Jospin in 1998 \citep{BurnettBonami2019}. I therefore treat 1986 as a near-null policy, take 1991--1997 as the window in which Switzerland is treated and France is not, and hold the 1998 French change back as a separate follow-up event.

I measure the proportion of profession and title nouns that take feminine marking for a female referent, counted strategy-agnostically: any feminine article, agreement, or noun form rather than one favoured ending. The two varieties differ a little in preferred form, Switzerland tolerating \mention{-esse} where France resists feminine endings longer \citep{Dawes2003}, but sharing the language spares the comparison the gulf between French, German, and Italian. Because that single count assumes the realizations are exchangeable signs of one change, I break it out by strategy, separating a distinct feminine noun form (\mention{autrice}) from a feminine article or agreement on an otherwise invariant noun (\mention{la ministre}); for the many epicene nouns the article and agreement are the only feminization signal, so a divergence between strategies is a warning, not a detail.

Female reference comes off independent cues, chiefly \mention{Madame} with a title, so the \mention{le}/\mention{la} alternation is identified without hand-coding. Two costs ride with that convenience. It narrows the outcome to a formal title-address context, so the estimand is feminine marking there, not across profession nouns at large. And \mention{Madame}-implies-female-referent is itself a classifier with an error rate; should \mention{Madame} usage drift across the window, that error drifts with it and lands in the measurement term $b$, the thing rung 3 makes its lesson.

Suisse romande's press is far smaller than France's, so the rarer titles land in thin cells. Partial pooling across outlets and titles within each variety steadies those sparse estimates, though for a two-period contrast the pre and post averaging already does most of that work, so the real payoff waits for the staggered design of rung 2 and its many group-time cells. The two varieties stay fixed effects so the contrast at issue is not pooled away, and the pooling stops short of the editorial layer the press adds, which stays on display.

Two cautions belong up front. The corpora are edited newspapers, so the design identifies a change in the edited written standard, not in spontaneous usage; use in the wider population stays a separate target, and for this phenomenon it may be out of reach by construction: the registers where female reference is cleanly conditioned are the edited and transcribed ones, and the spontaneous registers are exactly where neither the conditioning nor the transcription can be trusted.

Corpus frequency is also a population rate plus a composition term, so the usual parallel-trends assumption now has to hold twice over: the corpus adds parallel trends in the composition term to the standard requirement of parallel trends in the rate, rather than replacing it. Rung 1's target is the usage rate (Table~\ref{tab:ledger}), so the response to that second requirement is to reweight to a fixed composition, the policy-moved topic and profession mix included, since for a usage estimand that mix is nuisance, not effect. The one margin left untouched is the realization strategy itself, which is part of the outcome.

The two groups buy one sharp question, with a trap inside it. If the 1991 endorsement changed usage, Suisse romande should accelerate away from France after 1991 while the two ran together before. But an endorsement often ratifies a change already under way, so a pre-1991 divergence would mean 1991 is a waypoint on a curve, not its onset, and the later gap is no policy effect. The pre-1991 slopes come first: does the adoption rate change at 1991, or does a gap merely open? If instead both series rise together in calendar time throughout, the driver is a shared francophone shift in norms, and the contrast reflects that wave, not the intervention.

The threat this rung introduces is interference. Suisse romande reads French media, and both draw on the same wire service, so the control is not sealed off from the treated variety, and the estimand has to be named: the Swiss policy's direct effect, net of what crosses the border, or its total effect including diffusion. Agence France-Presse share and French-media dependence become exposure variables rather than nuisances to delete, shared-wire copy is held aside for a separate estimate, and the report states how much of each series it carries. A jump that lives only in shared material is diffusion, not a Swiss policy effect.

The procedure also runs on deliberately broken data, shuffling the treatment year to random dates and permuting the variety labels, to confirm it finds nothing where nothing should be. With only two varieties it's a falsification and calibration check, not powered inference: a clean panel offers no real randomization, so the permuted contrasts show the size of jump the pipeline manufactures by chance, against which the observed contrast is read. I do not present a permutation $p$-value as though the two varieties had been randomly assigned \citep{BertrandDufloMullainathan2004}.

Run the ladder of Section~\ref{sec:diagnostics} on this case. A divergence at 1991 that survives the composition-margin, shared-wire, and onset checks, with a breakdown value beyond any plausible ideology shift, would license a \emph{bounded estimate} of a policy effect on the edited written standard; a divergence leaning on an ambiguous check would license only a \emph{descriptive reading}; calendar-time co-movement with no break at 1991 would be read as a \emph{shared wave}, the francophone one; and a pre-1991 divergence, an effect carried by shared-wire or single-outlet copy, or a plausible ideology shift that overturns the estimate, would force \emph{not identified}. The routing is fixed before the estimate is seen, and a population claim is licensed by none of these on its own; it waits for the spoken and lightly edited series.

Rung 1 thus carries the whole apparatus in miniature, a stated estimand, a falsifiable parallel-trends assumption, an interference diagnostic, a randomization check, and the Section~\ref{sec:diagnostics} ladder applied, on the cleanest data the phenomenon offers. The rungs that follow keep the apparatus and the rule and add the complications cross-variety work cannot avoid.

\subsection{Rung 2: more polities, less exogeneity}

Adding varieties looks like a free improvement and is not. Québec endorsed feminine titles in 1979 \citep{oqlf1979} and Belgium by décret in 1993 \citep{belgique1993decret}; with Suisse romande (1991) and France (1998) that's four polities adopting at four dates. Treatment timing now varies, so the two-group contrast gives way to staggered estimators: group-time effects read against not-yet-treated varieties \citep{CallawaySantAnna2021}, after the standard warning that a pooled two-way fixed-effects estimate is a weighted average of two-by-two contrasts whose weights can go negative \citep{GoodmanBacon2021,deChaisemartin2020}, with event-study estimators robust to that contamination \citep{SunAbraham2021,BorusyakJaravelSpiess2024}. These estimators fix one thing, the negative weighting that heterogeneous effects induce; none of them touches selection into timing, the threat the next paragraphs turn to.

Worse, the not-yet-treated control isn't evenly available. France, last to adopt, can stand as the untreated comparison for Switzerland and Belgium, but its own 1998 change has no not-yet-treated peer inside the four, and Québec, treated since 1979, is no clean control for most of the window either. So the estimator that reads each cohort against the not-yet-treated has fewer clean comparisons than four polities suggest, the rung's lesson in miniature: more varieties bought not more identification but a thinner set of usable contrasts.

But four polities are not four independent experiments, and this rung exists to show why. The varieties are now whole countries, each with its own politics, press economy, and media regime that the two-group baseline never had to hold apart. The forms diverge as well: Québec writes \mention{auteure}, Belgium leans epicene, France resists, so rung 1's strategy-agnostic count now rests on treating these as exchangeable signs of one change, a premise to test rather than assume \citep{Dawes2003}.

Deeper still: the staggering isn't exogenous. It's the selection into treatment timing of Section~\ref{sec:threats}, now spread across cohorts: each polity adopts roughly when its own gender politics have already turned, and that turn moves usage too, so the not-yet-treated varieties are not resting on a flat counterfactual but riding the continental wave, the secular gender-ideology trend behind both adoption timing and usage, easily mistaken for the effect. The control is contaminated in the one dimension the design needs clean, and no estimator repairs it. \textcite{RambachanRoth2023} can bound how large a trends violation would have to be to overturn a result, but it cannot manufacture exogeneity where treatment timing is selected on the outcome's own trajectory; and with four clusters the large-sample approximations behind ordinary standard errors fail, since they need far more groups than four, leaving cluster-robust intervals anti-conservative and the permutation set tiny \citep{BertrandDufloMullainathan2004}.

What the staggering does buy is a sharper form of rung 1's question. If the policies move usage, the jumps should track each polity's own date; if a shared wave moves it, they should track calendar time whatever the date. The test is a partial filter, not a rescue: it separates one synchronized wave from policy-locked timing, but cannot tell a policy effect from heterogeneous local waves that each drive their own polity's date. France supplies the internal placebo: its 1986 circular went unenforced, on the government's own later admission, and should show no jump where 1998 does, so an effect that turns up at 1986 reflects something other than the policy \citep{BurnettBonami2019,Fujimura2005}. It's a negative control for one date, not a warrant for the parallel trends the other varieties still require.

The decision ladder carries over unchanged in form, with the staggered diagnostics slotted in. A cross-variety divergence aligned to policy dates, surviving the composition, shared-wire, and pre-trend checks, with breakdown values beyond a plausible wave, licenses a \emph{bounded estimate}; calendar-time alignment is read as the \emph{shared wave}; a failed France placebo, contamination revealed by the decomposition, or a plausible wave that overturns the estimate forces \emph{not identified}. More varieties raise the bar for a clean claim at least as often as they lower it.

\FloatBarrier
\subsection{What the simulations show, and where they break}

A corpus never reveals the population rate, so the one place to check that the design recovers the right answer is a simulation whose truth I set: the only corpus in the paper honest enough to reveal its $\pi$. On controlled data the claim ladder routes each canonical case to the verdict it should, and reweighting recovers the true effect when a composition shift rides an observed stratum: a naive contrast of $+0.13$ on a pure composition artifact returns to zero once the measured mix is held fixed, while drift outside the strata leaves both the naive and the reweighted estimate near $+0.15$, the case the design cannot catch (\texttt{rung1-recovery-sim.py}). That's a consistency check, not evidence the diagnostics have power on messy data.

As the scripts implement it, the ladder is a fixed, ordered sequence of gates on the reweighted series, short-circuiting at the first that fires. A pre-trend gate: a treated-minus-comparison pre-period slope beyond tolerance returns \emph{not identified}, the divergence predating the shock. A composition gate: a shift on a pre-specified reweightable margin triggers the reweighted estimand, and the naive contrast is disqualified where it diverges from the reweighted one beyond tolerance; only a shift on an unmeasured or non-reweightable margin returns \emph{not identified}. An interference gate: measured cross-border exposure beyond tolerance returns \emph{not identified}. A sensitivity gate: a reweighted DiD that's cleared the earlier gates but sits inside a band set at $1.5$ times the largest placebo-date contrast is a bounded null, an effect too small to separate from zero, while one that clears the band is a \emph{bounded estimate}. Telling that bounded null from a \emph{shared wave} is rung 2's calendar-time test, not a rung-1 gate; the \emph{descriptive} verdict of the full ladder, for a check that's ambiguous rather than passed or failed, is the human judgment the automated routine leaves aside. The tolerances are fixed before any run and listed in the code.

\begin{table}[t]
\centering
\footnotesize
\caption{Operating characteristics of the decision ladder on the rung-1 design, four hundred replications per scenario (\texttt{rung-failure-map.py}, seeded; mid-range rates carry a couple of points of Monte Carlo error). FP is a false effect claim, FN a missed real effect.}
\label{tab:failuremap}
\begin{tabular}{@{}l l r@{}}
\toprule
Scenario & Threat dialed up & Misrouting \\
\midrule
clean effect                     & none                            & 0\% FN \\
measured composition artifact    & composition, inside the strata  & 0\% FP \\
hidden drift, share $0.3$        & composition, outside the strata & 56\% FP \\
hidden drift, share $0.6$        & the same                        & 99\% FP \\
hidden drift, share $0.9$        & the same                        & 100\% FP \\
small effect, sparse ($n=40$)    & thin cells                      & 94\% FN \\
spillover, exposure measured     & interference                    & 0\% FP \\
spillover, exposure noisy        & interference, mismeasured       & 8\% FP \\
selection $\to$ wave (two-group) & timing on its own trajectory    & 100\% FP \\
\bottomrule
\end{tabular}
\end{table}

Each run draws cell proportions for two varieties over twenty years on a log-odds scale, routes the synthetic series through the whole ladder, and counts a verdict as a misroute when it disagrees with the truth I set; \enquote{hidden share} is the fraction of the post-period shift placed outside the strata the reweighting can see, and the sparse case shrinks the treated cells to forty tokens. So the failures get mapped (Table~\ref{tab:failuremap}), recording how often the ladder reaches the wrong verdict as each threat is turned up over four hundred replications.

Measured composition is reweighted away, but drift outside the reweighting strata is invisible, and the false-positive rate climbs from $56\%$ at a hidden share of $0.3$ to near-certainty by $0.6$. The sensitivity gate fails here too: the placebo benchmark it's built from sits in periods the hidden drift never touches, so the bar stays low while the contrast jumps. The simulation makes Section~\ref{sec:diagnostics}'s limit quantitative: the bound catches a violation that leaves a footprint in the benchmark windows, and only that.

Thin cells destroy power: a real effect in a sparse variety is missed in $94$ runs of a hundred. Interference is caught when exposure is measured well and leaks an $8\%$ false-positive rate when it isn't.

And selection into treatment timing, a control on its own trajectory, is observationally identical to a real effect in a two-group design. Rung 2's calendar-time test catches only one version, a single synchronized wave; heterogeneous local waves, each moving its own polity's timing, also produce date-aligned jumps and survive it. For a phenomenon driven that way the honest modal verdict is descriptive or not identified.

Read together, the simulations are sanity checks and a failure map, not a certificate that the method works. They mark the conditions under which a corpus DiD earns a bounded claim, and the conditions under which the verdict is descriptive or not identified.

\subsection{Rung 3: one institution, moving measurement}

The third rung changes the comparison axis rather than the estimator. Trilingual Switzerland holds the polity and its federal policy fixed while language varies, so a contrast across the French-, German-, and Italian-speaking regions asks whether a change is language-internal (a francophone wave shared with France) or polity-driven (a Swiss federal endorsement felt in all three). Because the federal policy reaches every region, none is an untreated control for it: the cross-language contrast separates a language-internal trajectory from a polity-wide one, but a clean policy effect would need an outside, language-matched comparison or a triple difference, which is why this rung is a design variant, not a third two-group DiD.

One could reach for a triple difference, a DiD of two DiDs built to net out a second confound, French-Switzerland against French-France crossed with the within-Switzerland language regions, but the German and Italian arms have no francophone-wave-matched control, so the third difference wouldn't subtract the wave cleanly, and the rung stays a probe of the measurement link rather than an identification device. The Federal Assembly's \emph{Bulletin officiel} is the natural home: one institution publishing in all three languages, digitized from 1891 and carrying full verbatim debates from 1971, so the 1991 window sits well inside the usable record and offers the within-country, language-varying counterfactual rungs 1 and 2 cannot.

But the register carries a cost, and the cost is the rung's lesson. Parliamentary address is squarely on topic, \mention{Madame la ministre} against \mention{Madame le ministre} is a real chamber-floor variable, though the official record is a normalized transcription, not speech, with its own house style. If the transcription service adopts feminine forms on some date, the observed frequency jumps with no change in what members said: the measurement link of Figure~\ref{fig:threat-dag}, from latent rate to observed $f$, becomes the dominant path, a policy shock on the corpus rather than on usage. The diagnostic is to date the transcription convention independently and treat its change as a rival event. The speaker pool drifts too, since the share of women members rises across the period and with it the referent base the outcome conditions on.

% Optional future extensions: Italian single-shock (Sabatini 1987); Croatian/Serbian breakdown.
