\section{Conclusion}
\label{sec:conclusion}

The discipline this paper asks for comes down to one refusal: not to read a corpus frequency as a population rate. Once that line holds, the rest follows. The observed contrast is the effect on the rate plus whatever the corpus filter contributed, so a causal claim needs the target population, the composition, the measurement, and the units of variation and inference pulled apart, with parallel trends asked of the rate and the filter both.

The payoff is not usually a clean effect. A corpus DiD is at its best when it pushes a causal claim downward, when the diagnostics turn an apparent effect into a shared wave, a measurement shock at a coding change, or a composition artifact, or show that it can't be identified from the series at all. Each is an answer worth reporting, not a result to bury. But honesty cuts both ways: a downgrade earned against the evidence is a finding, while the same verdict forced only by thin data is a missed effect mistaken for a success, which is why the failure map counts false negatives as carefully as false positives.

Before-and-after corpus data can carry a causal claim, but only under conditions worth stating plainly. The treatment has to be a named act at a named date; the comparison has to make a counterfactual plausible rather than decorate the design; composition and measurement have to be held fixed on pre-committed non-policy margins, or shown not to move; and the result has to survive a sensitivity bound larger than any confound the setting plausibly supplies. Where those hold, the claim is a bounded estimate scoped to the register actually measured, not a fact about the language at large.

More often they don't all hold, and then the output is smaller. A handful of identifying units leaves inference thin; treatment timing that tracks the same trend it's meant to measure can't be cleanly separated from an effect; composition or measurement drift that goes unmeasured survives every reweighting. None of that counts as a defect of the method; it's the method reporting what the data can and cannot support, the report most worth having.

The feminization case carried these points without resting on any of them: a place to watch the apparatus work and break. The same apparatus travels to other shocks and other languages, as far as the corpora reach. When the next speakers arrive, a new medium spreads, a border moves, or a decree lands, and a corpus frequency shifts with it, the discipline turns the question from whether the line moved to what moved: the rate, the filter, the measurement, or only the warrant for the word \mention{cause}. The apparatus offers no way to get more causal claims out of corpus data. What it offers is a way to make the ones worth trusting, and to know, before publishing, which those are.
